Algebraic Groups\vspace{-1ex}

### 2019-12-04 Algebraic Geometry/Algebraic-Geometry.tex
update

We took some time to talk about limits as (right) adjoints to the constant diagram functor. Hadn't seen that before.
We concluded with a bunch of cateory theory, which was really fun. Ind and Pro categories.
\subsection{More examples}
Let either
\begin{itemize}
	\item $X$ and $Y$ be spaces
	\item $h_Y:U\mapsto h_Y(U)$ is a sheaf on $X$ for all open $U$ in $X$.
\end{itemize}

>>>>

We took some time to talk about limits as (right) adjoints to the constant diagram functor. Hadn't seen that before.
We concluded with a bunch of cateory theory, which was really fun. Ind and Pro categories.
\subsection{More examples}
\begin{itemize}
	\item Let $X$ and $Y$ be spaces
	\item $h_Y:U\mapsto h_Y(U)$ is a sheaf on $X$ for all open $U$ in $X$.
	\item If $X$ is a manifold and $V\to X$ is a vector bundle over $X$, then you get a variety of sheaves (discrete, cts, flat, holomorphic sections).
\end{itemize}

\section{December 4th, 2019}
Let $\calC$ be a category (think of $\Set$).
\subsection{More Sheafiness}
\begin{defn}
	A \textbf{presheaf} with values in $\calC$ on $X$ is an $\calF$ such that 
	\begin{itemize}
		\item $\forall U\subseteq X$ (open) we have $\calF(U)\in\calC$
		\item $\forall V\subseteq U\subseteq X$, we get a restriction map $\rho_{VU}:F(U)\to F(V)$ which composes with other restrictions.
		\item $\rho_{UU}=\id_{F(U)}$
	\end{itemize}
\end{defn}
\begin{rmk}
	Of course this is all equivalent to saying a presheaf is a functor 
	\[\calF:\mathbf{Open}(X)^{op}\to \calC.\]
\end{rmk}

Now let's say that $\calC$ has products (including the empty product/final object). Then definitely the sheaf condition makes sense. 
We get the coordinate kind of definition if we imaging $\calC$ is $\Set$ (or at least concrete).
\begin{defn}
	A \textbf{sheaf} is a presheaf that satisfies the sheaf condition for all open coverings of all open sets.
\end{defn}
Some examples:
\begin{itemize}
	\item $\O=h_{\A1}$
	\item $h_Y$ is a sheaf on $X$ as we saw above.
\end{itemize}
\begin{rmk}
	There are two categories:
	\[\PreSh=\Func(\Open(X)\op,\calC)\]
	and a full subcaltegory $\Sh$ of sheaves.
\end{rmk}
\subsection{Stalks}
\begin{defn}
	Given a $p\in X$W and $\calC\in \PreSh(X)$, the \textbf{stalk of $\calF$ at $p$} is
	\[\calF_p=\colim_{p\in U}\calF(U)\]
\end{defn}
\begin{rmk}
	If $\calC$ is $\Set$, then we can realize this as $\sqcup_{p\in U}\calF(U)$ modulo the relation that two maps are the same
	if they agree on a small neighborhood of $p$.
\end{rmk}
As a matter of notation, we say that if $p\in V$ and $f\in\calF(V)$, then there is a value $f_p\in \calF_p$,
called the \textbf{germ of $f$ at $p$.}

\begin{thm}[Basic]
	A map of sheaves $\calF\to \calG$ is an isomorphism (resp. mono/epi) if and only if for all $p\in X$ the induced map $\calF_p\to\calG_p$ is an iso (resp. mono, epi).
\end{thm}
\begin{prob}
	For next time: Classify all (pre)sheaves on the space with two points $x$ and $y$ with topology $\{\varnothing,\{x\},\{x,y\}\}$. Also think about monos, epis, and isos of (pre)sheaves).
\end{prob}

### 2019-12-04 Algebraic Groups/Algebraic-Groups.tex
update

	Now let $A\subseteq \End_k(V)$ generated by $G$. Let $M=V$ (which is assumed to be simple as above). Then by letting $B=\End_A(V)=k$, so by the above 
	result, $\End_k(V)=A$ so $A$ is simple! Then let $I=\langle g-I_V:g\in G\rangle$, a two-sided ideal of $A$, so is either zero (in which case we are done) or is all of $A$.

\end{prf}

>>>>

	Now let $A\subseteq \End_k(V)$ generated by $G$. Let $M=V$ (which is assumed to be simple as above). Then by letting $B=\End_A(V)=k$, so by the above 
	result, $\End_k(V)=A$ so $A$ is simple! Then let $I=\langle g-I_V:g\in G\rangle$, a two-sided ideal of $A$, so is either zero (in which case we are done) or is all of $A$.
\end{prf}

\section{December 4th, 2019}
Recall the Kolchin FPT from last time. 
\subsection{More unipotent stuff}
\begin{cor}
	If $V$ is a finite dimensional vector space over $k$ and $G\to\GL(V)$ is an abstract group homomorphism whose image consists of unipotent elements, then 
	there exists a basis of $V$ such that $G\subseteq U_n\subseteq\GL_n$ where $U_n$ is the group of upper-triangular matrices with ones on the diagonal.
\end{cor}
\begin{cor}
	If $G$ is asmooth linear algebraic group over $k=\bar k$ such that $G(k)$ consists of unipotents, then $G\subseteq U_n$ as a closed subgroup.
\end{cor}
Examples include $\Ga$ (included as $(\begin{smallmatrix}
	1&a\\0&1
\end{smallmatrix})$) which is unipotent. In characteristic $p$, $\bbZ/p$ is unipotent. There is also some weird $\alpha_p$ stuff.
\begin{defn}
	A linear algebraic group $G$ over $k$ is unipotent if for all nonzero $G$-representations, $V^G\ne 0$.
\end{defn}
\begin{rmk}
	This is equivalent if we require $V$ to be finite dimensional over $k$.
\end{rmk}
\begin{prop}
	$G$ is unipotent if and only if for all finite dimensional representations $\rho:G\to \GL(V)$, there exists a basis 
	such that $\rho(G)\subseteq U_n$.
\end{prop}
\begin{prf}
	Choose a filtration of $V$:
	\[0=V_0\subset V_1\subset\cdots\subset V_n=V\]
	with all quotients $W_i\eqdef V_i/V_{i-1}$ simple. But then since $W_i^G\ne 0$ by the definition of unipotence, $W_i^G=W_i$.
	So then $W_i$ is the trivial representation and apparently that finishes the forward direction.

	The reverse direction is easy (pick a basis and $v_1$ is fixed by $\rho(G)$).
\end{prf}
\begin{thm}
	Let $G$ be a linear algebraic group over $k$. Then $G$ is unipotent if and only if $G$ is closed in $U_n$ for some $n$.
\end{thm}
\begin{lem}
	If $G$ is unipotent, so is any subgroup, quotient, or extension.
\end{lem}
\subsection{Induced representation}
Let $H\to G$ be a linear algebraic group homomorphism. Then we have a functor 
\[\operatorname{Ind}_H^G:\Rep(H)\to\Rep(G)\]
that is defined via 
\[\operatorname{Ind}_H^G(V)=\Mor_{\Sch/k}^H(G,\bbA(V))=\{G\to \bbA(V)|\text{$H$ invariant}\}=(\Gamma(G)\otimes_k V)^H\]
Notice that the inclusion functor $f:\mathbf{B}H\to\mathbf{B}G$ gives us two maps, $f_\ast$ (induction) and $f^\ast$ (forgetful).
